Place of death in NRS COVID-19 deaths

### 2021-04-16 covid_nrs_place_of_death_visualisations.Rmd
adding cause of death breakdown

@@

```{r}
(figure_annual_deaths_by_cause <-
  total_deaths_cause_of_death %>%
  filter(year == 2020 & cause_of_death!="All" & !place_of_death %in% c("Other","All")) %>%
  # pivot_longer(cols = c(deaths, annual_deaths_mean), names_to = "deaths_type", values_to = "deaths") %>%
  ggplot() +
  # geom_col(aes(x = cause_of_death, fill = cause_of_death, y = deaths, group = deaths_type, linetype = deaths_type), position = "dodge", colour = "grey10") +
  geom_col(aes(x = cause_of_death, fill = cause_of_death, y = deaths), colour = "grey10") +
  # geom_point(aes(x = cause_of_death, y = annual_deaths_mean), colour = "grey10") +
  geom_line(aes(x = cause_of_death, y = annual_deaths_mean), colour = "grey10") +
  scale_fill_viridis_d() +
  facet_wrap(~place_of_death) +
  NULL
)
```

@@

```{r, fig.width=12, fig.height=9, warning=FALSE}
(figure_deaths_at_home_by_cause <-
  merged_deaths_cause_of_death %>%
  filter(str_starts(place_of_death,"Home")) %>%
  filter(cause_of_death != "All") %>%
  # filter(!place_of_death %in% c("All","Other")) %>%
  pivot_longer(cols = c(deaths_mean, number_of_deaths), names_to = "deaths_type", values_to = "number_of_deaths") %>%
  filter(!(deaths_type=="deaths_mean"&cause_of_death=="COVID-19")) %>%  # remove historical average for covid, it's 0 anyway
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home"))) %>%  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)
  ggplot(aes(x = date_w_c)) +
  # annotate(geom = "rect", xmin = compute_start_date_from_week_number(week_number = 1, year_number = 2021), xmax = compute_start_date_from_week_number(week_number = weeks_available_2021, year_number = 2021), ymin = -Inf, ymax = Inf, fill = "grey90", alpha = 0.2) +  # shade area - geom_rect doesn't work here
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "8 weeks", limits = c(earliest_date, most_recent_date_available_2021)) +
  geom_line(aes(y = number_of_deaths, colour = cause_of_death, linetype = deaths_type)) +
  scale_colour_viridis_d() +
  scale_linetype_manual(labels = c("Historic mean 2015-2019","Deaths since 2020"), values = c("dashed","solid")) +
  # geom_line(aes(y = number_of_deaths  , colour = deaths_type)) +
  # geom_line(aes(y = deaths_mean), linetype = "dashed", colour="grey20") +
  facet_wrap(cause_of_death~place_of_death, scales = "free_y") +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
    plot.margin = margin(5, 15, 5, 5),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
  ) +
  labs(x = "Date w/c",
       y = "Weekly Deaths (N)",
       subtitle = date_range_text,
       caption = paste0(c("Note that y-axis scale varies across panels!",death_registrations_text,removed_other_category_text,source_nrs_text), collapse="\n")
       ) +
  NULL
)

>>>>

## Table of annual deaths by cause

```{r}
table_deaths_by_cause_2020 <-
  total_deaths_cause_of_death %>%
  select(-matches("min|max")) %>%
  # filter(year == 2020 & cause_of_death!="All" & !place_of_death %in% c("Other","All")) %>%
  filter(year == 2020 & cause_of_death!="All") %>%
  group_by(place_of_death) %>%
  mutate(
    proportion_annual_deaths = deaths / sum(deaths),
    historical_proportion_annual_deaths = annual_deaths_mean / sum(annual_deaths_mean)
    ) %>%
  ungroup

table_deaths_by_cause_2020 %>%
  mutate(across(.cols = matches("ratio|prop"), .fns = ~scales::percent(.x, accuracy=0.1))) %>%
  select(Period = year, "Place of death" = place_of_death, "Cause of death" = cause_of_death, "Annual deaths" = deaths, "Proportion of annual deaths" = proportion_annual_deaths, "Historic average" = annual_deaths_mean, "Historic proportion" = historical_proportion_annual_deaths, "Ratio 2020 to historic" = ratio_annual_deaths_to_historical_mean) %>%
  write_csv(file = paste0(dir_outputs,"/table_causes_of_death_2020.csv"))
```

@@

```{r}
(figure_annual_deaths_by_cause <-
  total_deaths_cause_of_death %>%
  filter(year == 2020 & cause_of_death!="All" & !place_of_death %in% c("Other","All")) %>%
  # pivot_longer(cols = c(deaths, annual_deaths_mean), names_to = "deaths_type", values_to = "deaths") %>%
  ggplot() +
  # geom_col(aes(x = cause_of_death, fill = cause_of_death, y = deaths, group = deaths_type, linetype = deaths_type), position = "dodge", colour = "grey10") +
  geom_col(aes(x = cause_of_death, fill = cause_of_death, y = deaths), colour = "grey10") +
  # geom_point(aes(x = cause_of_death, y = annual_deaths_mean), colour = "grey10") +
  # geom_line(aes(x = cause_of_death, y = annual_deaths_mean), colour = "grey10") +
  scale_fill_viridis_d() +
  facet_wrap(~place_of_death) +
  NULL
)
```

@@

```{r, fig.width=12, fig.height=9, warning=FALSE}
(figure_deaths_at_home_by_cause <-
  merged_deaths_cause_of_death %>%
  filter(str_starts(place_of_death,"Home")) %>%
  filter(cause_of_death != "All") %>%
  # filter(!place_of_death %in% c("All","Other")) %>%
  pivot_longer(cols = c(deaths_mean, number_of_deaths), names_to = "deaths_type", values_to = "number_of_deaths") %>%
  filter(!(deaths_type=="deaths_mean"&cause_of_death=="COVID-19")) %>%  # remove historical average for covid, it's 0 anyway
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home"))) %>%  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)
  ggplot(aes(x = date_w_c)) +
  # annotate(geom = "rect", xmin = compute_start_date_from_week_number(week_number = 1, year_number = 2021), xmax = compute_start_date_from_week_number(week_number = weeks_available_2021, year_number = 2021), ymin = -Inf, ymax = Inf, fill = "grey90", alpha = 0.2) +  # shade area - geom_rect doesn't work here
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "8 weeks", limits = c(earliest_date, most_recent_date_available_2021)) +
  # geom_line(aes(y = number_of_deaths, colour = cause_of_death, linetype = deaths_type)) +
  geom_line(aes(y = number_of_deaths, linetype = deaths_type)) +
  scale_colour_viridis_d() +
  scale_linetype_manual(labels = c("Historic mean 2015-2019","Deaths since 2020"), values = c("dashed","solid")) +
  # geom_line(aes(y = number_of_deaths  , colour = deaths_type)) +
  # geom_line(aes(y = deaths_mean), linetype = "dashed", colour="grey20") +
  facet_wrap(cause_of_death~place_of_death, scales = "free_y") +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
    plot.margin = margin(5, 15, 5, 5),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
  ) +
  labs(x = "Date w/c",
       y = "Weekly Deaths (N)",
       subtitle = date_range_text,
       caption = paste0(c("Note that y-axis scale varies across panels!",death_registrations_text,source_nrs_text), collapse="\n")
       ) +
  NULL
)
